Towards Comparable Active Learning

### 2023-04-28 doc/al_benchmark_paper.tex
TopV2 datasets

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with an extensive experiments section. \\
{\color{red} TODO} We include in this table any methodological paper that experiments on at least two domains. \\
{\color{red} TODO} Define AL scenarios (really hard)
\begin{table}[h]
	\centering
	\begin{tabular}{l | c c c c c}
		Paper & \# Datasets & Domains & Scenarios & Oracle & RL Alg. \\
		\hline
		Beck et al. \cite{beck2021effective} & 4 & 1 & 3 & - & - \\
		Hu et al. \cite{hu2021towards} & 5 & 2 & 1 & - & - \\
		Li et al. \cite{li2022empirical} & 5 & 1 & 1 & - & - \\
		\textbf{Ours} & 5 & 2 & 2 & \checkmark & -
	\end{tabular}
	\caption{Comparison of our benchmark with the existing literature}
	\label{tab:benchmark_comparison}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Problem Description / Delineation}
{\color{red} Version: 1.0}\\
Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$. \\
Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
\begin{align*}
	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
	\quad\quad \text{\tiny (equivariant in the second argument)}
\end{align*}
% which is equivariant in the second argument, 
called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal:
\begin{align*}
	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
	% \operatorname{avg} \{
	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
	\ell(y, \hat y(x)) 
	\\
	\text{with }
	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
	\\ 
	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
	\quad b\in 1{:}B
\end{align*}
(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
Since combinatorial problem of finding the optimal subset $\mathcal{D}_\text{train}$ is computational not feasible, we apply two relaxations to the problem. 
Firstly, we allow sequential construction of $\mathcal{D}_\text{train}$ and secondly, we limit the choice of data points in every iteration to a random sample of size $\tau$. \\ [2mm]

\textbf{Pool-based AL with single instances (non-batch setting)}\\
To construct the active learning setting, we suppress the labels $y_i$ of $\mathcal{D}$ to form the unlabeled pool $\mathcal{U} := u_i \hspace{0.5mm}; \hspace{1mm} i := 1 \ldots N$ and form and initial labeled pool $\mathcal{L}$ by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label. 
The result of this so-called "seeding" process is $\mathcal{L} := (u_i, y_i) \hspace{0.5mm}; \hspace{1mm} i := 1 \ldots k*C$. \\
Active learning is defined as sequentially removing single instances $u^{(i)} \in \mathcal{U}^{(i)}\hspace{1mm}; \hspace{2mm} \mathcal{U}^{(i+1)} := \mathcal{U}^{(i)} \setminus´ \{u^{(i)}\}$, recovering their label $y^{(i)}$ and adding them to the labeled pool $\mathcal{L}^{(i+1)} := \mathcal{L}^{(i)} \cup (u^{(i)}, y^{(i)})$ until a fixed budget $B$ is exhausted $i := 1 \ldots B$.
After each added instance the classification model is retrained according to section \ref{sec:training_the_classifier} and its performance is measured on the held-out test set $\mathcal{D}^{\text{test}}$.
The quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(y_{test}, \hat y_i(x_{test}))
\end{equation}
Where $\hat y_i$ is the retrained classification model after the i-\textit{th} instance was selected. \\ [1mm]
%
\textbf{Framing AL as RL}\\
We define the active learning process as an adapted reinforcement learning loop $(S, A, \tau, \Omega, \omega)$ where an environment iteratively will expose a state $s \in S$ to an agent $\Omega$, which will choose actions $a \in A$.
For each iteration $i$ the environment samples a subset of size $\tau$ of unlabeled instances $u^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$, constructs the state $s^{(i)} := \omega(u^{(i)})$ and presents it to the agent to select an action $a^{(i)} := \Omega(s^{(i)})$.
The action $a^{(i)}$ is the index of the selected instance in $u^{(i)}$ out of all possible indices $A := [1 \ldots \tau]$.
This process is repeated $B$ times $i := [1 \ldots B]$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Splice Embedded}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\linewidth]{img/eval_splice_embedded}
	\caption{Results for all algorithms on the pre-encoded Splice dataset}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.8\linewidth]{img/eval_splice_embedded_cd}
\caption{Critical Difference Diagram for Splice where every restart of the algorithm is one sample for the Wilcoxon-Holm method}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.8\linewidth]{img/eval_macro}
\caption{Critical Difference Diagram for Splice, DNA, USPS and Cifar10}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with an extensive experiments section. \\
{\color{red} TODO} We include in this table any methodological paper that experiments on at least two domains. \\
{\color{red} TODO} Define AL scenarios (really hard)
\begin{table}[h]
	\centering
	\begin{tabular}{l | c c c c c}
		Paper & \# Datasets & Domains & Scenarios & Oracle & RL Alg. \\
		\hline
		Beck et al. \cite{beck2021effective} & 4 & 1 & 3 & - & - \\
		Hu et al. \cite{hu2021towards} & 5 & 2 & 1 & - & - \\
		Li et al. \cite{li2022empirical} & 5 & 1 & 1 & - & - \\
		Zhou et al. \cite{zhou2021towards} & 3 & 2 & 1 & \checkmark & - \\
		\textbf{Ours} & 5 & 2 & 2 & \checkmark & -
	\end{tabular}
	\caption{Comparison of our benchmark with the existing literature}
	\label{tab:benchmark_comparison}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Problem Description}
{\color{red} Version: 1.0}\\
Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$. \\
Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
\begin{align*}
	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
	\quad\quad \text{\tiny (equivariant in the second argument)}
\end{align*}
% which is equivariant in the second argument, 
called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal:
\begin{align*}
	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
	% \operatorname{avg} \{
	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
	\ell(y, \hat y(x)) 
	\\
	\text{with }
	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
	\\ 
	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
	\quad b\in 1{:}B
\end{align*}
(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
Since combinatorial problem of finding the optimal subset $\mathcal{D}_\text{train}$ is computational not feasible, we apply two relaxations to the problem. 
Firstly, we allow sequential construction of $\mathcal{D}_\text{train}$ and secondly, we limit the choice of data points in every iteration to a random sample of size $\tau$. \\ [2mm]
%
To construct the active learning setting, we suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$. \\ [2mm]

Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(y_{test}, \hat y_i(x_{test}))
\end{equation}
Where $\hat y_i$ is the retrained classification model after the i-\textit{th} instance was selected. \\ [1mm]
%
\textbf{Framing AL as RL}\\
We define the active learning process as an adapted reinforcement learning loop $(S, A, \tau, \Omega, \omega)$ where an environment iteratively will expose a state $s \in S$ to an agent $\Omega$, which will choose actions $a \in A$.
For each iteration $i$ the environment samples a subset of size $\tau$ of unlabeled instances $u^{(i)} \underset{\tau}{\sim} \mathcal{U}^{(i)}$, constructs the state $s^{(i)} := \omega(u^{(i)})$ and presents it to the agent to select an action $a^{(i)} := \Omega(s^{(i)})$.
The action $a^{(i)}$ is the index of the selected instance in $u^{(i)}$ out of all possible indices $A := [1 \ldots \tau]$.
This process is repeated $B$ times $i := [1 \ldots B]$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Splice Embedded}

\begin{table}[]
	\begin{tabular}{l|lll}
		& Splice        & DNA           & USPS          \\
		\hline
		Oracle          & 0.835 +- 0.01 & 0.879 +- 0.01 & 0.870 +- 0.01 \\
		SAL\_23\_01\_16 & 0.808 +- 0.01 & 0.857 +- 0.02 & 0.865 +- 0.01 \\
		Coreset\_Greedy & 0.808 +- 0.01 & 0.853 +- 0.02 & 0.849 +- 0.01 \\
		MarginScore     & 0.806 +- 0.01 & 0.855 +- 0.02 & 0.864 +- 0.01 \\
		ShannonEntropy  & 0.808 +- 0.01 & 0.855 +- 0.02 & 0.864 +- 0.01 \\
		RandomAgent     & 0.797 +- 0.01 & 0.823 +- 0.02 & 0.831 +- 0.01
	\end{tabular}
\end{table}
%
\begin{table}[]
	\begin{tabular}{l|ll}
		& Cifar10 & FashionMnist \\
		\hline
		Oracle          & 0.718 +- 0.01  & 0.721 +- 0.01       \\
		SAL\_23\_01\_16 & 0.660 +- 0.01  & 0.667 +- 0.01       \\
		Coreset\_Greedy & 0.668 +- 0.01  & 0.674 +- 0.01       \\
		MarginScore     & 0.674 +- 0.01  & 0.677 +- 0.01       \\
		ShannonEntropy  & 0.671 +- 0.01  & 0.680 +- 0.01       \\
		RandomAgent     & 0.632 +- 0.02  & 0.633 +- 0.01      
	\end{tabular}
\end{table}
%
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\linewidth]{img/eval_splice_embedded}
	\caption{Results for all algorithms on the pre-encoded Splice dataset}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.8\linewidth]{img/eval_splice_embedded_cd}
\caption{Critical Difference Diagram for Splice where every restart of the algorithm is one sample for the Wilcoxon-Holm method}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.8\linewidth]{img/eval_macro}
\caption{Critical Difference Diagram for Splice, DNA, USPS and Cifar10}
\end{figure}
